---
output:
  pdf_document: default
  html_document: default
---
DATA EXPLORATION

```{r}
library('tidyverse')
library('readr')

rawData <- data.frame(read_csv('https://intro-datascience.s3.us-east-2.amazonaws.com/Resort01.csv'))
str(rawData) 

summary(rawData)
#View(rawData)
```


```{r}
#Check for NA values in all columns

table(is.na(rawData$IsCanceled))
table(is.na(rawData$LeadTime))
table(is.na(rawData$StaysInWeekendNights))
table(is.na(rawData$StaysInWeekNights))
table(is.na(rawData$Adults))

table(is.na(rawData$Children))
table(is.na(rawData$Babies))
table(is.na(rawData$Meal))
table(is.na(rawData$Country))
table(rawData$Country)#NULL values


table(is.na(rawData$MarketSegment))
table(is.na(rawData$IsRepeatedGuest))
table(is.na(rawData$PreviousCancellations))
table(is.na(rawData$PreviousBookingsNotCanceled))
table(is.na(rawData$ReservedRoomType))
table(is.na(rawData$AssignedRoomType))
table(is.na(rawData$BookingChanges))
table(is.na(rawData$DepositType))
table(is.na(rawData$CustomerType))
table(is.na(rawData$RequiredCarParkingSpaces))
table(is.na(rawData$TotalOfSpecialRequests))

```
DATA CLEANING

```{r}
#Data Cleaning


#Data cleansing Meal and Country column, change lead time column name
#to days since booking
str(rawData)

table(rawData$Meal)
#1169 undefined in Meal columN

table(rawData$Country)
table(rawData$MarketSegment)

table(rawData$ReservedRoomType)
table(rawData$AssignedRoomType)
table(rawData$DepositType)
table(rawData$CustomerType)
hist(rawData$LeadTime,breaks=5)
```

```{r}
#Cleaning
#Filter out rows where the number of adults is less than one  
newData1 <- rawData[select(rawData,Adults) >= 1,]



#To standardize the dataset values, we replaced rows that contain “Undefined” 
#values for the “meal” column with “SC”, as they mean the same thing 
#(mentioned by the contributor of the dataset).
newData1$Meal[newData1$Meal=="Undefined"]<-"SC"
table(newData1$Meal)

#The Country attribute has NULL values that are filtered out
newData2 <- newData1[select(newData1,Country)!="NULL",]
table(newData2$Country)

#Rename column LeadTime to make it easier to understand the context of the 
#column

colnames(newData2)[2] <- "DaysSinceBooking"


rough <- newData2[newData2$StaysInWeekendNights==0 & 
              newData2$StaysInWeekNights==0 ,  ]

newData3 <- anti_join(newData2,rough)


table(newData3$Country)


hotelData <- newData3
str(hotelData)
table(hotelData$ModifiedCountryCode)
#View(hotelData)
```

DATA TRANSFORMATION
```{r}

newData3$ModifiedCountryCode<- newData3$Country
newData3$ModifiedCountryCode[newData3$ModifiedCountryCode=="CN"] <- "CHN"
table(hotelData$ModifiedCountryCode)

#New detractor column created
hotelData$Detractor <- as.factor(hotelData$PreviousCancellations>12)
hotelData <- newData3
str(hotelData)
table(hotelData$ModifiedCountryCode)
```


```{r}
#install.packages("viridis")
library(viridis)

#Percentage of Bookings vs Cancellation Status
#Here we are figuring out the persentage of canccellations in the whole dataset
#about 28%

counts3 <- table(hotelData$IsCanceled)
library(RColorBrewer)
coul <- brewer.pal(5, "Pastel1") 
barplot(prop.table(counts3) * 100, main="Percentage of Bookings vs Cancellation Status",
  xlab="Hotel Cancellations (0- not canceled, 1- canceled)",ylab="Percentage of bookings",col=coul,
  legend = rownames(counts3), beside=TRUE)


```

```{r}
#Boxplot and histogram of the numerical variables 
str(hotelData)

#DaysSinceBooking
g1 <- ggplot(hotelData, aes(x=factor(IsCanceled), y=DaysSinceBooking, fill=IsCanceled)) + geom_boxplot(fill="lightblue")
g1

ggplot(data = hotelData )+
       aes(x = as.factor(IsCanceled), y = DaysSinceBooking) + 
    geom_bar(stat = 'identity', position = 'dodge',fill=rgb(1,0.8,0.8,0.7))
```


```{r}
#StaysInWeekendNights
g2 <- ggplot(hotelData, aes(x=factor(IsCanceled), y=StaysInWeekendNights, fill=IsCanceled)) + geom_boxplot(fill="lightblue")
g2
ggplot(data = hotelData )+
       aes(x = as.factor(IsCanceled), y = StaysInWeekendNights) + 
    geom_bar(stat = 'identity', position = 'dodge',fill=rgb(1,0.8,0.8,0.7))
```


```{r}
#StaysInWeekNights
g3 <- ggplot(hotelData, aes(x=factor(IsCanceled), y=StaysInWeekNights, fill=IsCanceled)) + geom_boxplot(fill="lightblue")
g3
ggplot(data=hotelData,aes(StaysInWeekNights)) +geom_histogram(binwidth=3)
```


```{r}
#Adults
g4 <- ggplot(hotelData, aes(x=factor(IsCanceled), y=Adults, fill=IsCanceled)) + geom_boxplot(fill="lightblue")
g4
ggplot(data = hotelData )+
       aes(x = as.factor(IsCanceled), y = Adults) + 
    geom_bar(stat = 'identity', position = 'dodge',fill=rgb(1,0.8,0.8,0.7))
```


```{r}
#PreviousCancellations
g5 <- ggplot(hotelData, aes(x=factor(IsCanceled), y=PreviousCancellations, fill=IsCanceled)) + geom_boxplot(fill="lightblue")
g5
ggplot(data = hotelData )+
       aes(x = as.factor(IsCanceled), y = PreviousBookingsNotCanceled) + 
    geom_bar(stat = 'identity', position = 'dodge',,fill=rgb(1,0.8,0.8,0.7))
```


```{r}
#PreviousBookingsNotCanceled
g6 <- ggplot(hotelData, aes(x=factor(IsCanceled), y=PreviousBookingsNotCanceled, fill=IsCanceled)) + geom_boxplot(fill="lightblue")
g6
ggplot(data = hotelData )+
       aes(x = as.factor(IsCanceled), y = PreviousBookingsNotCanceled) + 
    geom_bar(stat = 'identity', position = 'dodge',,fill=rgb(1,0.8,0.8,0.7))
```


```{r}
#Children
g7 <- ggplot(hotelData, aes(x=factor(IsCanceled), y=Children, fill=IsCanceled)) + geom_boxplot(fill="lightblue")
g7
ggplot(data = hotelData )+
       aes(x = as.factor(IsCanceled), y = Children) + 
    geom_bar(stat = 'identity', position = 'dodge',fill=rgb(1,0.8,0.8,0.7))
```


```{r}
#Babies
g8 <- ggplot(hotelData, aes(x=factor(IsCanceled), y=Babies, fill=IsCanceled)) + geom_boxplot(fill="lightblue")
g8

ggplot(data = hotelData )+
       aes(x = as.factor(IsCanceled), y = Babies) + 
    geom_bar(stat = 'identity', position = 'dodge',fill=rgb(1,0.8,0.8,0.7))
```


```{r}
#BookingChanges
g9 <- ggplot(hotelData, aes(x=factor(IsCanceled), y=BookingChanges, fill=IsCanceled)) + geom_boxplot(fill="lightblue")
g9

ggplot(data = hotelData )+
       aes(x = as.factor(IsCanceled), y = BookingChanges) + 
    geom_bar(stat = 'identity', position = 'dodge',fill=rgb(1,0.8,0.8,0.7))
```


```{r}
#RequiredCarParkingSpaces
g10 <- ggplot(hotelData, aes(x=factor(IsCanceled), y=RequiredCarParkingSpaces, fill=IsCanceled)) + geom_boxplot(fill="lightblue")
g10
ggplot(data = hotelData )+
       aes(x = as.factor(IsCanceled), y = RequiredCarParkingSpaces) + 
    geom_bar(stat = 'identity', position = 'dodge',fill=rgb(1,0.8,0.8,0.7))
```


```{r}
#TotalOfSpecialRequests
g11 <- ggplot(hotelData, aes(x=factor(IsCanceled), y=TotalOfSpecialRequests, fill=IsCanceled)) + geom_boxplot(fill="lightblue")
g11
ggplot(data = hotelData )+
       aes(x = as.factor(IsCanceled), y = TotalOfSpecialRequests) + 
    geom_bar(stat = 'identity', position = 'dodge',fill=rgb(1,0.8,0.8,0.7))

```

```{r}
#dividing data into subsets of cancelled and not cancelled

str(hotelData)
cancelData <- hotelData[hotelData$IsCanceled==1,]
notCancelData <- hotelData[hotelData$IsCanceled==0,]

```

```{r}
#Map visualization

#install.packages('rworldmap')
library('rworldmap')

cancelData$IsCanceled < -as.factor(cancelData$IsCanceled)
help(joinCountryData2Map)
sPDF <- joinCountryData2Map(hotelData,joinCode = "ISO3",nameJoinColumn = "ModifiedCountryCode",mapResolution = "coarse",nameCountryColumn = "ModifiedCountryCode",verbose=TRUE)

#CN and GIB country codes could not be mapped


mapParams <- mapCountryData( sPDF, nameColumnToPlot="IsCanceled",addLegend = FALSE)


#The above map gives a color graded country wise representation based on the number of days since booking incase of cancelled bookings 

par(mar=c(2,2,2,2))
worldmap <- joinCountryData2Map(cancelData, joinCode="ISO3", nameJoinColumn="Country",)
map<-mapCountryData(worldmap, nameColumnToPlot='DaysSinceBooking',  catMethod="logFixedWidth", colourPalette=c('white','green','red','yellow','blue','black','orange'))
```



```{r}

#Bar-plots
#Identifying trends between the different categories in categorical type data in Cancelled Bookings

#Meal type
ggplot(cancelData, aes(x=Meal)) +
  geom_bar(color="blue", fill=rgb(0.1,0.4,0.5,0.7) )


#Market Segment
ggplot(cancelData, aes(x=MarketSegment)) +
  geom_bar(color="blue", fill=rgb(0.1,0.4,0.5,0.7) )

#DepositType
ggplot(cancelData, aes(x=DepositType)) +
  geom_bar(color="blue", fill=rgb(0.1,0.4,0.5,0.7) )

#CustomerType
ggplot(cancelData, aes(x=CustomerType)) +
  geom_bar(color="blue", fill=rgb(0.1,0.4,0.5,0.7) )

#Assigned Room Types
counts6 <- table(cancelData$AssignedRoomType)
library(RColorBrewer)
coul <- brewer.pal(5, "Set2") 
barplot(prop.table(counts6) * 100, main="Percentage of Bookings vs AssignedRoomType",
  xlab="Assigned Room Types",ylab="Percentage of bookings",col=viridis(10),
  legend = rownames(counts6), beside=TRUE)


```

Market segment frequency in entire dataset
```{r}
library(ggplot2)



table1<-table(hotelData$MarketSegment) 
table1
barplot(table1,col=magma(6),main="Market Segment",xlab="Different Market Segments",ylab="Frequency",border="black",
        legend.text = rownames(table1),args.legend = list(x = "top"))

```

Not Cancelled/ Cancelled Rooms W.R.T Market Segment
```{r}


df9<-hotelData[,c('IsCanceled','MarketSegment')]
table2<-table(df9$IsCanceled,df9$MarketSegment)
barplot(table2,beside = T, col=c("#edf8b1","#7fcdbb"),main="Cancelation Status vs Market Segments",xlab="Different Market Segments",ylab="Frequency",border="black",
        cex.names =.6,legend.text =rownames(table2),args.legend = list(x = "top",title="IsCanceled"))

```

Market segment frequency wrt StaysInWeekendNights and StaysInWeekNights
could or could not be cancelled

```{r}
#Not Cancelled/ Cancelled Rooms W.R.T Meals

df10<-hotelData[,c('IsCanceled','Meal')]
table3<-table(df10$IsCanceled,df10$Meal)
barplot(table3,beside = T, col=c("#edf8b1","#7fcdbb"),main="Cancelation Status vs Meals",xlab="Meals",ylab="Frequency",border="black",
        cex.names =.8,legend.text =rownames(table3),args.legend = list(x = "top",title="IsCanceled"))

```

Cancelled Rooms market segment frequency wrt StaysInWeekendNights and StaysInWeekNights

```{r}
df3<-cancelData[,c('StaysInWeekendNights','StaysInWeekNights','MarketSegment')]
dfplot3<-pivot_longer(df3, -MarketSegment, names_to="variable", values_to="value")
ggplot(dfplot3,aes(x = MarketSegment,y = value)) + 
    geom_bar(aes(fill = variable),stat = "identity",position = "dodge") +ggtitle("Cancelled Rooms Market Segment for Stays during Weekday or Weekend")+ xlab("Cancelled Rooms Market Segments")+ylab("Frequency")

```

Meal frequency in entire Hotel data

```{r}
table4<-table(hotelData$Meal) 
table4
barplot(table4,col=viridis(4),main="Meals Frequency",xlab="Different Meals",ylab="Frequency",border="black",
        legend.text = rownames(table4),args.legend = list(x = "top"))

```

Not Cancelled/ Cancelled Rooms W.R.T Meals

```{r}
df10<-hotelData[,c('IsCanceled','Meal')]
table3<-table(df10$IsCanceled,df10$Meal)
barplot(table3,beside = T, col=magma(2),main="Not Cancelled/ Cancelled Rooms W.R.T Meals",xlab="Meals",ylab="Frequency",border="black",
        cex.names =.8,legend.text =rownames(table3),args.legend = list(x = "top",title="IsCanceled"))

```

Meal frequency wrt StaysInWeekendNights and StaysInWeekNights
could or could not be cancelled

```{r}
df4<-hotelData[,c('StaysInWeekendNights','StaysInWeekNights','Meal')]
dfplot4<-pivot_longer(df4, -Meal, names_to="variable", values_to="value")
ggplot(dfplot4,aes(x = Meal,y = value)) + 
    geom_bar(aes(fill = variable),stat = "identity",position = "dodge") +ggtitle("Meals for Stays during Weekday or Weekend")+ xlab("Meals")+ylab("Frequency")
```

Cancelled Rooms Meals frequency wrt StaysInWeekendNights and StaysInWeekNights
```{r}
dim(cancelData)
df5<-cancelData[,c('StaysInWeekendNights','StaysInWeekNights','Meal')]
dfplot5<-pivot_longer(df5, -Meal, names_to="Legend", values_to="value")
ggplot(dfplot5,aes(x = Meal,y = value)) +   geom_bar(aes(fill = Legend),stat = "identity",position = "dodge") +ggtitle("StaysInWeekendNights & StaysInWeekNights vs Meals in Canceled Bookings")+ xlab("Cancelled Rooms Meals")+ylab("Frequency")


```
DepositType frequency 

```{r}

table6<-table(hotelData$DepositType) 
table6
barplot(table6,col=rainbow(3),main="Deposit Type Frequency",xlab="Deposit Types",ylab="Frequency",border="black",
        legend.text = rownames(table6),args.legend = list(x = "top"))
```

Not Cancelled/ Cancelled Rooms W.R.T Deposit type

```{r}
df11<-hotelData[,c('IsCanceled','DepositType')]
table5<-table(df11$IsCanceled,df11$DepositType)
barplot(table5,beside = T, col=rainbow(2),main="Not Cancelled/ Cancelled Rooms W.R.T Deposit Type",xlab="Deposit Type",ylab="Frequency",border="black",
        cex.names =.8,legend.text =rownames(table5),args.legend = list(x = "top",title="IsCanceled"))

```

DepositType frequency wrt StaysInWeekendNights and StaysInWeekNights
could or could not be cancelled

```{r}

df7<-hotelData[,c('StaysInWeekendNights','StaysInWeekNights','DepositType')]
dfplot7<-pivot_longer(df7, -DepositType, names_to="variable", values_to="value")
ggplot(dfplot7,aes(x = DepositType,y = value)) + 
    geom_bar(aes(fill = variable),stat = "identity",position = "dodge") +ggtitle("Deposit Types for Stays during Weekday or Weekend")+ xlab("Deposit Types")+ylab("Frequency")

```

Cancelled DepositType frequency wrt StaysInWeekendNights and StaysInWeekNights

```{r}
df8<-cancelData[,c('StaysInWeekendNights','StaysInWeekNights','DepositType')]
dfplot8<-pivot_longer(df8, -DepositType, names_to="variable", values_to="value")
ggplot(dfplot8,aes(x = DepositType,y = value)) + 
    geom_bar(aes(fill = variable),stat = "identity",position = "dodge") +ggtitle("Cancelled Deposit Types for Stays during Weekday or Weekend")+ xlab("Cancelled Rooms Deposit Types")+ylab("Frequency")


```

RequiredCarParkingSpaces in Cancelled and Not Cancelled Data

```{r}
table(cancelData$RequiredCarParkingSpaces)
table(hotelData$RequiredCarParkingSpaces)

Graph11<-hotelData[,c('IsCanceled','RequiredCarParkingSpaces')]
Graph11<-table(Graph11$IsCanceled,Graph11$RequiredCarParkingSpaces)
barplot(Graph11,beside = T, col=(c("#edf8b1","#7fcdbb")),main="Required Parking Spaces on Cancelled and Not Cancelled Data",xlab="Required Parking Spaces",ylab="Frequency",border="black",
        cex.names =.8,legend.text =rownames(Graph11),args.legend = list(x = "top",title="IsCanceled"))

```


NPS (Detractor and Promotor)
```{r}

table(hotelData$PreviousCancellations)

vectorforNPS<- hotelData$PreviousCancellations
totalNPS<-length(vectorforNPS)

hotelData$Detractor<- "FALSE"
hotelData$Detractor[hotelData$PreviousCancellations>12] <- "TRUE"
table(hotelData$Detractor)

vectorDetractor<- as.vector(hotelData$Detractor=="TRUE")
numDetractor<-sum(vectorDetractor)
numDetractor

hotelData$Promotor<- "FALSE"
hotelData$Promotor[hotelData$Detractor == "FALSE"] <- "TRUE"
table(hotelData$Promotor)

vectorPromotor<- as.vector(hotelData$Promotor=="TRUE")
numPromotor<-sum(vectorPromotor)
numPromotor

nps<- (numPromotor/totalNPS - numDetractor/totalNPS) *100
nps

```
**MODELING**


Logistic Regression
```{r}
#Logistic Regression


modelLog <- glm(IsCanceled ~DaysSinceBooking+StaysInWeekendNights+StaysInWeekendNights+StaysInWeekNights+PreviousCancellations+PreviousBookingsNotCanceled+BookingChanges+TotalOfSpecialRequests+MarketSegment+CustomerType+RequiredCarParkingSpaces+TotalOfSpecialRequests+DepositType+ReservedRoomType+Meal,family=binomial(link='logit'),data=hotelData)

summary(modelLog)


```

Support Vector Machines

(1)
```{r}
#Support Vector Machines

mergedDF <- hotelData
str(mergedDF)

mergedDF$IsCanceled <- as.factor(mergedDF$IsCanceled)
#mergedDF$Meal <- as.factor(mergedDF$Meal)
#mergedDF$MarketSegment <- as.factor(mergedDF$MarketSegment)
str(mergedDF)
library(caret)
library(kernlab)
library(arules)
library(arulesViz)
library(tidyverse)
set.seed(111)


mergedDF2 <- mergedDF %>% select(IsCanceled,DaysSinceBooking,StaysInWeekendNights,Meal,MarketSegment,IsRepeatedGuest,PreviousCancellations,DepositType,CustomerType,RequiredCarParkingSpaces,TotalOfSpecialRequests,BookingChanges)

trainList <- createDataPartition(y=mergedDF2$IsCanceled, p=.70, list=FALSE)

trainData <- mergedDF2 [trainList,]
testData <- mergedDF2 [-trainList,]

svm.model <- train(IsCanceled ~ ., data = trainData, method="svmRadial", trControl=trainControl(method="none"), preProcess=c("center", "scale"))
svm.model

predict_Val <- predict(svm.model, newdata = trainData)
#predict_Val
table(predict_Val,trainData$IsCanceled)
conf_Matrix <- confusionMatrix(predict_Val, trainData$IsCanceled)

fourfoldplot(table(predict_Val,trainData$IsCanceled))
```



Association Rule Mining

```{r}
##Association Rule Mining

library(arules)
library(arulesViz)

#summary(mergedDF3)
mergedDFnew <- mergedDF2
str(mergedDFnew)
mergedDFnew$IsCanceled <- as.factor(mergedDFnew$IsCanceled)
mergedDFnew$DaysSinceBooking <- as.factor(mergedDFnew$DaysSinceBooking)
mergedDFnew$StaysInWeekendNights <- as.factor(mergedDFnew$StaysInWeekendNights)
#mergedDFnew$StaysInWeekNights <- as.factor(mergedDFnew$StaysInWeekNights)
#mergedDFnew$Adults <- as.factor(mergedDFnew$Adults)
#mergedDFnew$Children <- as.factor(mergedDFnew$Children)
mergedDFnew$Meal <- as.factor(mergedDFnew$Meal)
mergedDFnew$MarketSegment <- as.factor(mergedDFnew$MarketSegment)
mergedDFnew$IsRepeatedGuest <- as.factor(mergedDFnew$IsRepeatedGuest)
mergedDFnew$PreviousCancellations <- as.factor(mergedDFnew$PreviousCancellations)
mergedDFnew$DepositType <- as.factor(mergedDFnew$DepositType)
mergedDFnew$CustomerType <- as.factor(mergedDFnew$CustomerType)
mergedDFnew$RequiredCarParkingSpaces <- as.factor(mergedDFnew$RequiredCarParkingSpaces)
mergedDFnew$TotalOfSpecialRequests <- as.factor(mergedDFnew$TotalOfSpecialRequests)
mergedDFnew$RequiredCarParkingSpaces <- as.factor(mergedDFnew$RequiredCarParkingSpaces)
mergedDFnew$BookingChanges <- as.factor(mergedDFnew$BookingChanges)




set.seed(111)
mergedDFnew <- as(mergedDFnew,"transactions")

itemFreq <- itemFrequency(mergedDFnew)

str(itemFreq)
sortedData <- sort(itemFreq)
itemFreqPlot <- itemFrequencyPlot(mergedDFnew, topN=20)
 
rules <- apriori(mergedDFnew, parameter=list(supp=0.042, conf=0.9), control=list(verbose=F), appearance=list(default="lhs",rhs=("IsCanceled=1")))
inspect(rules)
plot(rules,method="paracoord")
```


